---
title: "Introduction to Data Science Assignment 04"
subtitle: "Food Security"
author: "Daniela Nagar dn386, Ronin McCobb rcm125"
format: 
  pdf:
    code-line-numbers: true
execute: 
  warning: false
urlcolor: blue
editor_options: 
  chunk_output_type: console
---

\vspace{0.1in}
\newpage
```{r}
library(tidyverse)
library(ipumsr)
library(srvyr)
library(haven)

ddi <- read_ipums_ddi("cps_00004.xml")
data <- read_ipums_micro(ddi)
```

\newpage
```{r}
#Data Visualization 1: Number of Respondents per Food Security Score in the US
data |>
  filter(FSRAWSCRA < 11 & FSRAWSCRA > 0 ) |>
  ggplot(aes(x=FSRAWSCRA)) +
  geom_bar(width = 1, fill = "blue") +
   labs(
      title = "Number of Respondents per Food Security Scores 
      in the US",
      caption = "Sarah Flood, Miriam King, Renae Rodgers, Steven Ruggles,
J. Robert Warren, Daniel Backman, Annie Chen, Grace Cooper, Stephanie Richards,
Megan Schouweiler, and Michael Westberry. IPUMS CPS Version 12.0 [dataset]. 
Minneapolis, MN: IPUMS, 2024. https://doi.org/10.18128/D030.V12.0"
    ) +
  xlab("Score") +
  ylab("Number of Respondents")

```

#Written interpretation: To start, we looked at the number of respondents across all states who have every food security score. Food security scores span from 1-10. Respondents with larger scores are more food insecure. The graph reflects what we would expect by trending downward as scores grow higher, indicating that fewer people are lacking in all food security measures but many are lacking in at least one measure.
